The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2018-08-30 JStatSoft_forestinventory_master-concordance.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 108 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 116:%
1 98 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 215:%
124 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 220:%
1 67 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 288:%
130 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 293:%
1 104 1 1 4 1 1 1 2 1 0 2 1 3 0 1 2 2 1 1 3 14 0 1 2 16 1 1 5 7 0 1 2 4 1 1 2 24 0 1 2 7 1 1 2 4 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 1 2 8 0 1 2 139 1 1 7 6 0 1 1 26 0 1 2 3 1 1 2 13 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 7 1 1 6 5 0 1 1 8 0 1 2 7 1 1 2 4 0 1 2 1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 6 5 0 1 1 8 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
136}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
1 105 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 1 1 1 6 8 0 1 2 2 1 1 2 27 0 1 2 150 1 1 3 2 0 1 1 3 0 1 2 4 1 1 8 10 0 1 2 1 1 1 4 1 1 1 2 12 0 1 2 4 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 6 1 1 7 9 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1237:%
138 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1242:%
1 35 1 1 2 11 0 1 2 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1294:%
144 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1299:%
1 12 1 1 2 4 0 1 2 4 1 1 6 8 0 1 2 4 1 1 6 8 0 1 2 11 1 1 8 16 1 1 7 9 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 3 5 0 1 2 28 1 1 3 5 0 1 2 27 1 1 3 5 0 1 2 25 1 1 3 5 0 1 2 18 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1548:%
150 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1553:%
1 9 1 1 4 6 0 1 2 3 1 1 6 8 0 1 2 2 1 1 7 9 0 1 2 2 1 1 7 9 0 1 2 2 1 1 8 10 0 1 2 4 1 1 5 7 0 1 2 6 1 1 2 4 0 1 2 2 1 1 2 26 0 1 2 9 1 1 4 1 1 1 2 11 0 1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1748:%
156 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1753:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1765:%
162 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1770:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1775:%
168 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 105 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
1 64 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 178:%
121 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 183:%
1 67 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 251:%
127 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 256:%
1 2 1 1 4 107 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
0 1 2 4 1 1 2 24 0 1 2 30 1 1 3 5 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 %
1 2 8 0 1 2 130 1 1 5 7 0 1 2 2 1 1 2 27 0 1 2 3 1 1 2 13 0 1 2 5 1 %
1 5 4 0 1 1 8 0 1 2 7 1 1 5 7 0 1 2 2 1 1 2 9 0 1 2 7 1 1 3 5 0 1 2 %
1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 5 7 0 1 2 2 1 1 2 9 0 %
1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
133}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
1 2 1 1 4 91 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 29 0 1 2 %
102 1 1 3 2 0 1 1 3 0 1 2 4 1 1 5 7 0 1 2 3 1 1 2 12 0 1 2 4 1 1 5 7 %
0 1 2 2 1 1 2 12 0 1 2 6 1 1 5 7 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1169:%
135 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1174:%
1 2 1 1 4 38 1 1 2 11 0 1 2 3 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1232:%
141 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1237:%
1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
7 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 2 4 0 1 2 27 1 1 2 4 %
0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1472:%
147 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1477:%
1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
0 1 2 2 1 1 6 8 0 1 2 4 1 1 4 6 0 1 2 6 1 1 2 4 0 1 2 40 1 1 2 11 0 %
1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1666:%
153 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1671:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1683:%
159 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1688:%
1 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1693:%
165 11 1}

### 2018-08-30 JStatSoft_forestinventory_master.rnw
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable evidence-based information to assess the state and development of forests over time.  They typically consist of a random sample of plot locations in the forest that are assessed individually by field crews. Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision. With respect to this objective, the application of multi-phase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient. While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multi-phase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata. This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in the \proglang{R} environment.}

@@

\begin{document}\sloppy

>>>>

%% an abstract and keywords
\Abstract{
Forest inventories provide reliable evidence-based information to assess the state and development of forests over time.  They typically consist of a random sample of plot locations in the forest that are assessed individually by field crews. Due to the high costs of these terrestrial campaigns, remote sensing information available in high quantity and low costs is frequently incorporated in the estimation process in order to reduce inventory costs or improve estimation precision. With respect to this objective, the application of multiphase forest inventory methods (e.g., double- and triple-sampling regression estimators) has proved to be efficient. While these methods have been successfully applied in practice, the availability of open-source software has been rare if not non-existent. The \proglang{R} package \pkg{forestinventory} provides a comprehensive set of global and small area regression estimators for multiphase forest inventories under simple and cluster sampling. The implemented methods have been demonstrated in various scientific studies ranging from small to large scale forest inventories, and can be used for post-stratification, regression and regression within strata. This article summarizes the mathematical theory of this family of design-based estimators and demonstrates their application in the \proglang{R} environment.}

@@

\begin{document}\sloppy
\SweaveOpts{concordance=TRUE}

### 2018-08-30 revision/reviews.tex
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

@@

@@

The paper is rather long which can be a good or bad thing. I think that the 2 and 3 phase estimators are so similar that it may be sufficient to present, for example, SAE only for the 2 phase estimator and global estimation only for the 3 phase estimator. The same is true for the exhaustive and non-exhaustive estimators. Describing just once that in all cases both types are possible (as done for cluster sampling), would make the text a lot easier to read. Also section 5 on CIs could be considerably shortened. Needs to be a section on its own right? Sections 6 and 7 are very interesting but the R code under 7.1 could probably be condensed a lot.\\

@@

@@

@@

@@

@@

@@

\note{\textit{Andi}: There is a slight inconsistency in our article here: Our package says 'multiphase', but here we used 'multi-phase'. However, should we then also change 'two-phase' to 'twophase'?}

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

	\answerfin{Whereas we agree that they might be slightly more readable without the n2 fractions, the decision to include them was based on the fact that this is how the formulas were originally presented in Mandallaz's papers. We feel that the marginal loss in readability is offset by perfect continuity with the original notation.}
	% --------------------------- %

@@

	\answerfin{It has been added that the models require intercept terms.}
	% --------------------------- %

@@

The paper is rather long which can be a good or bad thing. I think that the 2 and 3 phase estimators are so similar that it may be sufficient to present, for example, SAE only for the 2 phase estimator and global estimation only for the 3 phase estimator. The same is true for the exhaustive and non-exhaustive estimators. Describing just once that in all cases both types are possible (as done for cluster sampling), would make the text a lot easier to read. \\
\\
\answerfin{Different versions of this article were written that attempted to shorten it in ways just like the review mentioned.  However, we ultimately decided that the purpose of this article is to bridge the gap between likely users of the software (e.g. forest engineers and technical practitioners) and the mathematical details that might be beyond their skillset.  For this reason, we decided to ensure that every function call was connected explicitly to the underlying formula that it applies and a citation is given for those interested in further technical details.  We believe this also enables the reader to write their own routines using the functions which can be used to error check our routines. Considering that the current length is not excessively long compared to other JSS papers, we strongly prefer to not to shorten sections merely for the sake of making the article shorter.} \\

Also section 5 on CIs could be considerably shortened. Needs to be a section on its own right? Sections 6 and 7 are very interesting but the R code under 7.1 could probably be condensed a lot.\\
\\
\answerfin{Reviewer 1 made a very valid criticism about designed-based confidence intervals when the sampling fraction is too small that we would like to include in the paper and Section 5 is the most logical place to put it.} \\

@@

\answerfin{When the package was orginally beta tested at the Swiss Federal Institute of Forest, Snow and Landscape Management on likely operational users as well as forest engineers, the most contentious area for confusion was the calculation of explanatory variables in the exhaustive case.  The reviewer appears to have confirmed that confusion in the previous paragraph, which is explained in the section on calculating explanatory variables.  It is explictly stated that the exact exhaustive mean vector can only be calculated for explanatory variables based on means (note that polygon layers are means of binary variables). It has now been explicitly added as an example that order statistics are not included in this category with the reviewers example given.  It is already explicitly stated that the size of the support should ideally be chosen to achieve a best possible explanatory power of the regression model on page 9 and a citation has been added for the interested reader.} \\

@@

\answerfin{The population doesn't need to be tessellated and can be allowed to overlap or not use all the wall-to-wall map in the infinite population approach. This is already mentioned on page 9.  Section 2.1 discusses the Infinite Population Approach.  Some brief discussion comparing to the finite is included, but as this package is about software written for the infinite not finite approach, we prefer not to delve too deeply into a topic that makes little practical difference.} \\

@@

\answerfin{This is true, but it is already stated many times that the synthetic estimators will usually have a smaller variance than estimators incorporating the regression model uncertainties, but at the cost of a potential bias making them seem deceptively optimistic (see pages 5, 13 (twice), 14, 16 (this is explained here using a concrete example), 21, 26 and 30)} \\

@@

\answerfin{Removing outliers is more of a common practice in model-dependent estimation, which is beyond the scope of this article.  You should never remove outliers in design-based estimation unless you can defend the belief that the resulting (and self-inflicted) non-response is missing at random.  The observed value of the the response/explanatory variable is fixed at each $x$ so deleting an observation effectively means the plot at x has been removed from the forest.  For temporal mismatches this is pretty much indefensible because the outliers likely represent harvesting after the aerial measurement.  We initially included a discussion comparing design-based with model-dependent estimation but we ultimately decided that the paper has more focus if we concentrate only on explaining how to use this software under the design-based infinite population approach.  In any case, many citations are currently included that already discuss these issues.} \\

@@

\answerfin{A mention of sample plot boundary adjustments is now included in Section 2.1 with a citation.} \\

@@

\answerfin{Multiphase is now used consistently to match package's spelling}
% --------------------------- %

@@

\answerfin{Good point.  Has been changed accordingly.}
% --------------------------- %

@@

\answerfin{That suggestion only can be done if the plots form a perfect tesselation.  Circular units, the most common in NFIs, can't do this.  The rest of the sentence actually says "with respect to the forest area".  This is true.  Points can fill any forested area even if it is a fractal whereas a counter example forest area can always be given for any plot shape you put forward.  Extending your tesselation beyond the forest edge is possible, but this increases the weights for boundary cases which likely increases the variance.   Many of the cited articles in the infinite population approach already discuss this but the focus for this article is not a comparison of the finite and infinite approaches, but a presentation of a software package written under the infinite approach.}
% --------------------------- %

@@

\answerfin{The g-weight is the name of a variance estimator for the point estimate in eq.2 as opposed to another possible variance estimator here named the external estimator.  The origin of the g-weight name is more obvious in the citation provided in the same paragraph.  The g-weight is analogous to a hat matrix and can be extracted from the more compact form that was included in the article.  It the spirit of not being too verbose we chose the compact form. }
% --------------------------- %

@@

\answerfin{In this context "a biased estimate" is an estimate that is produced from a biased estimator.  It would be a bit tedious to explain every time that an given estimate was produced from a biased estimator.}
% --------------------------- %

@@

\answerfin{The "survey" package is in the finite population approach (aka list sampling).  It is mentioned on page 2.}
% --------------------------- %

@@

\answerfin{It is elegant in the sense that the SAE estimator is unbiased merely by modifying the design matrix.  It is not necessary to refit every small area with an indicator, but it is  how the package implemented the estimator.}
% --------------------------- %

@@

\answerfin{It says that "the formulas look similar to the synthetic estimators", because they are exactly same.  They merely have a modified design matrix.}
% --------------------------- %

@@

\answerfin{Change has been implemented.}
% --------------------------- %

### 2018-08-30 rnw/Analysis and Visualization.Rnw
formula for rel.effiency corrected
This is the 2nd batch of changes including the section files.  The 1st batch and the second batch are commited separately but reverted to only the first batch.

As multi-phase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multi-phase global or small area estimate compared to its respective one-phase estimate. The function takes an \code{esttable} object as input and returns a summary of which multi-phase method and estimator performed best using the precision from the one-phase estimator as a baseline. If the \code{esttable} object contains more than one multi-phase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation. The argument \code{pref.vartype} can be used to define what type of variance (g-weight or external) should be used for the comparison. Synthetic estimates (SYNTH and PSYNTH estimator) are not considered for the comparison under the default setting (\code{exclude.synth = TRUE}) since they usually have a much smaller variance at the cost of a potential bias.

@@

% Interpretation: (gain and rel.eff)
The function call returns a data frame containing the one-phase variance (\code{var_onephase}) and the variance of the best performing multi-phase estimator (\code{var_multiphase}). The multi-phase estimation procedure is again specified in the \code{method} and \code{estimator} column. The last two columns quantify the potential benefit of the multi-phase estimation. The \code{gain} is the reduction (if its value is positive) in variance when applying the multi-phase as alternative to the one-phase estimation. For example, it is indicated that the two-phase extended PSYNTH estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure. The column \code{rel.eff} specifies the relative efficiency which is defined as the ratio between the one-phase variance and the multi-phase variance:

\begin{equation}\label{eq:releff}
  rel.eff_{[\%]}=\frac{\hat{\var}_{onephase}(\hat{Y})}{\hat{\var}_{multiphase}(\hat{Y})}\cdot100
  \end{equation}

The relative efficiency can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multi-phase estimator. For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the two-phase EXTPSYNTH estimator. If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multi-phase estimation for forest inventories.

>>>>

As multiphase estimation techniques are primary intended to increase estimation precision, the function \code{mphase.gain()} can be applied to quantify the potential benefit of a multiphase global or small area estimate compared to its respective one-phase estimate. The function takes an \code{esttable} object as input and returns a summary of which multiphase method and estimator performed best using the precision from the one-phase estimator as a baseline. If the \code{esttable} object contains more than one multiphase estimation object, \code{mphase.gain()} identifies the one with the smallest variance and compares it to the \code{onephase} estimation. The argument \code{pref.vartype} can be used to define what type of variance (g-weight or external) should be used for the comparison. Synthetic estimates (SYNTH and PSYNTH estimator) are not considered for the comparison under the default setting (\code{exclude.synth = TRUE}) since they usually have a much smaller variance at the cost of a potential bias.

@@

% Interpretation: (gain and rel.eff)
The function call returns a data frame containing the one-phase variance (\code{var_onephase}) and the variance of the best performing multiphase estimator (\code{var_multiphase}). The multiphase estimation procedure is again specified in the \code{method} and \code{estimator} column. The last two columns quantify the potential benefit of the multiphase estimation. The \code{gain} is the reduction (if its value is positive) in variance when applying the multiphase as alternative to the one-phase estimation. For example, it is indicated that the two-phase extended PSYNTH estimation procedure for small area \code{"B"} leads to a 67.9 \% reduction in variance compared to the one-phase procedure. The column \code{rel.eff} specifies the relative efficiency which is defined as the ratio between the one-phase variance and the multiphase variance:

\begin{equation}\label{eq:releff}
  rel.eff=\frac{\hat{\var}_{onephase}(\hat{Y})}{\hat{\var}_{multiphase}(\hat{Y})}
  \end{equation}

The relative efficiency can be interpreted as the relative sample size of the one-phase estimator needed to achieve the variance of the multiphase estimator. For small area \code{"B"} we can thus see that we would have to increase the terrestrial sample size by factor 3 in the one-phase approach in order to get the same estimation precision as the two-phase EXTPSYNTH estimator. If the average costs for a terrestrial sample plot survey are known, the relative efficiency can thus be a simple means of quantifying the financial benefit of using multiphase estimation for forest inventories.
